---
title: "Bellabeat Case Study"
author: "Bishop Sally"
date: "2023-06-01"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

# **Introduction** 

This is my breakdown of my Bellabeat Case study capstone project, which is part of my Google Data Analytics Certificate. This breakdown will give general insights to the analyst team at Bellabeat to make informed business decisions that can help the company in the future
The method I used to analyze Bellabeat was breaking everything down by using:

ASK 
PREPARE
PROCESS
ANALYZE
SHARE
ACT

## **About Bellabeat**

[Bellabeat](https://bellabeat.com/?gclid=CjwKCAjwg-GjBhBnEiwAMUvNW8dU3gg_7xVeMp_WV-1Acp6kc_57Kj7kPTRdLJCLFy2Cq3PRWln7cRoCep0QAvD_BwE) is a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

# **Ask**  

In the *Ask* section we should ask some key questions to find the problem and how we can solve them. The key questions that need to be asked to gain further insight are:
  **1.** What are the trends of smart device usuage?
  **2.** How can these trends apply to Bellabeat Customers?
  **3.**  How can these trends help influence Bellabeat Marketing Strategy 

**Business Task:** The questions that need to be asked will lead us to analyze smart device usage to gain insight into how consumers use non-Bellabeat smart devices. After we gain the insights we need we will then select one Bellabeat product that we can apply to these insights.

**How I will report the information:**
1. Give a clear summary of the business task
2. Adding a description of all data sources used
3. Documenting any cleaning or manipulation of data
4. Providing a summary of my analysis
5. Supporting visualizations and key findings
6. Your top high-level content recommendations based on your analysis

**Identifying Key Stakeholders:**
  *__Primary StakeHolders:__
  - **Urška Sršen:** Bellabeat’s cofounder and Chief Creative Officer
  - **Sando Mur:** Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team
  Secondary Stakeholders:
  *__Secondary Stakeholders:__
  - **Bellabeat marketing analytics team:** A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat’s marketing strategy

# **Prepare**
 In the prepare section we find, download, and store the data in a secured folder. 

* __What data was gathered:__
  [*Kaggle Dataset](https://www.kaggle.com/datasets/arashnic/fitbit) 
  *	This data set was generated by respondents to a distributed survey via [Amazon Mechanical Turk](https://zenodo.org/record/53894#.YMoUpnVKiP9) between 03.12.2016-05.12.2016. Thirty eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring.
* __Data Summary__:
  - 33 Fitbit users agreed to submit personal data while using their Fitbit Trackers.The data tracked contained different factors for each data set
  - There is a total of 18 csv files that were used for the [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit). Out of the 18, I used 8 out of the 18 data sets to help me find a solution. These data sets included: *daily_activity*, *daily_calories*, *daily_intensities*, *daily_steps*, *hourly_calories*, *hourly_steps*,*sleep_day*, and *weight_info*.
  - The data was generated between March 12 2016 through May 12 2016 
  - Source: Furberg, R., Brinton, J., Keating, M., & Ortiz, A. (2016). Crowd-sourced Fitbit data sets 03.12.2016-05.12.2016 [Data set]. Zenodo. https://doi.org/10.5281/zenodo.53894
  - [Creative Commons Atrribution 4.0 International](https://creativecommons.org/licenses/by/4.0/legalcode)

## **Data Reliability**
- 33 participants is considered a sufficient sample size, but a larger sample size would be more efficient based on the market size of the fitness tracker market as a whole. 
- The participants are also anonymous so it does not include specific demographic data to avoid any biases. 
- The data tracked activity and correlation but it does not state any specific features being utilized and tracked during this study.
- The data was generated between March and May 2016, so this data can be considered old and not being up to date with newer trends and features of Fitness Trackers

## **Conclusion:** I think overall this data is sufficient enough to get a general consensus of the Fitness tracker market, but a larger and more current dataset would be a good alternative in the future.

# **Process**

After securing the data now we can load, view, and clean the data to make sure the data is valid enough to complete the remaining steps

```{r}
# Installing necessary packages
options(repos = "https://cran.rstudio.com/")  # Set the CRAN mirror

install.packages("tidyverse")
install.packages("ggplot2")
install.packages("janitor")

```

```{r}
#Loading necessary packages 

library(tidyverse)
library(ggplot2)
library(janitor)
library(readr)
```
## **Importing Data**

```{r}
#Loading csv files into data sets 

daily_activity <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
View(daily_activity)

daily_calories <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
View(daily_calories)

daily_intensities <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/dailyIntensities_merged.csv")
View(daily_intensities)

daily_steps <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/dailySteps_merged.csv")
View(daily_steps)

sleep_day <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
View(sleep_day)

weight_info <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
View(weight_info)

hourly_steps <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/hourlySteps_merged.csv")
View(hourly_steps)

hourly_Calories <- read_csv("C:/Users/bsall/OneDrive/Documents/Fitabase Data 4.12.16-5.12.16/hourlyCalories_merged.csv")
View(hourly_Calories)

```

```{r}
# Removing unnecessary columns

daily_activity = subset(daily_activity,select = -c(LoggedActivitiesDistance, VeryActiveDistance, ModeratelyActiveDistance, LightActiveDistance, SedentaryActiveDistance))

daily_intensities = subset(daily_intensities,select = -c(SedentaryActiveDistance, LightActiveDistance, ModeratelyActiveDistance, VeryActiveDistance))

```


```{r}

# Finding null or missing values in my datasets

is.null(daily_activity)
is.null(daily_calories)
is.null(daily_intensities)
is.null(daily_steps)
is.null (sleep_day)
is.null(weight_info)
is.null(hourly_Calories)
is.null(hourly_steps)

```

```{r}
#Viewing the columns to check for any inconsistent data 

head(daily_activity)
str(daily_activity)
n_distinct(sleep_day$Id)
n_distinct(daily_activity$Id)

```

```{r}
is_empty(daily_activity)
is_empty(daily_calories)
is_empty(daily_intensities)
is_empty(daily_steps)
is_empty(sleep_day)
is_empty(weight_info)
is_empty(hourly_Calories)
is_empty(hourly_steps)

```

```{r}
# Updating and reformatting the date columns to get a more desired format 

daily_activity$ActivityDate <- as.Date(daily_activity$ActivityDate, "%m/%d/%y")
daily_calories$ActivityDay <- as.Date(daily_calories$ActivityDay, "%m/%d/%y")
daily_intensities$ActivityDay <- as.Date(daily_intensities$ActivityDay, "%m/%d/%y")
daily_steps$ActivityDay <- as.Date(daily_steps$ActivityDay, "%m/%d/%y")
sleep_day$SleepDay <- as.Date(strptime(sleep_day$SleepDay, "%m/%d/%Y"))
hourly_Calories$ActivityHour <- as.Date(hourly_Calories$ActivityHour, "%m/%d/%y")
hourly_steps$ActivityHour <- as.Date(hourly_steps$ActivityHour, "%m/%d/%y")
weight_info$Date <- as.Date(weight_info$Date, "%m/%d/%y")

```

```{r}
# Checking for the new date format 

str(daily_activity)
str(daily_calories)
str(daily_steps)
str(sleep_day)
str(daily_intensities)
str(hourly_Calories)
str(hourly_steps)
str(weight_info)

```
```{r}
# merging data sets: sleep_day and hourly_steps to create *combined_bars*

combined_bars <- merge(sleep_day, hourly_steps, by = c('Id'))

# viewing the newly merged data 

view(combined_bars)
head(combined_bars)


```
```{r}
# Merging the data sets: daily_activity and daily_steps to create *record_activity*

record_activity <- merge(daily_activity, daily_steps, by = c('Id'))

#Viewing the newly merged data set

view(record_activity)
head(record_activity)

```

```{r}
# Mutating all the time based columns to create a total minutes column that combines the times together 

total_intensities <- daily_intensities %>%
  mutate(daily_intensities = rowSums(select(., SedentaryMinutes, LightlyActiveMinutes, FairlyActiveMinutes, VeryActiveMinutes), na.rm = TRUE))

```

```{r}
#This string adds the calories column from the daily_calories data set into the total_intensities data set 

total_intensities$Calories <- daily_calories$Calories

#This changes the colname of daily_intensities into *TotalMinutes* to help organize the colnames when entering the data

names(total_intensities)[names(total_intensities) == "daily_intensities"] <- "TotalMinutes"

```
```{r}
# Removes all falsely reported data sets that were presented in the IsManualReport column 
weight_info <- subset(weight_info, IsManualReport == TRUE)

```

```{r}
#Summarizing weight of the selected columns and merged the data into the new variable *weight_summary* 

weight_summary <- weight_info %>%
  group_by(Date) %>%
  summarise(AverageWeight = mean(WeightPounds))

glimpse(weight_summary$AverageWeight)
glimpse(weight_info$WeightPounds)
glimpse(weight_info$Id)

```


# **Analyze**
After the data is cleaned now we can analyze the data so we can get an idea of how we are going to create and share the data 

```{r}
# Getting an overall summary of the data present in the *daily_activity* data set before I begin visualizing 
daily_activity %>%
  select(TotalSteps,
         TotalDistance, TrackerDistance, VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes, Calories) %>%
  summary()

```
```{r}
#Summarizing *sleep* and *Time* based columns in the **sleep_day** variable 
sleep_day %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary() 

```

```{r}
#Summarizing the *ActivityDate* and  *StepTotal* columns in the **record_activity** variable 

record_activity %>% 
  select(ActivityDate, StepTotal) %>%
  summary()

```


```{r}
#Summarizing the *TotalMinutes* and *Calories* columns in the **total_intensities** variable 

total_intensities %>%
  select(TotalMinutes,Calories) %>%
  summary()

```

```{r}
#Summarizing the *ActivityDate* and *StepTotal* columns in the **record_activity** variable 

record_activity %>% 
  select(ActivityDate, StepTotal) %>%
  summary()

```

```{r}
#Summarizing the *Date* and *AverageWeight* columns in the **weight_summary** variable 

weight_summary %>%
  select(Date,AverageWeight) %>%
  summary()

```

# **Share**

Now that we analyzed the different data sets now we can share the insights gathered to bring visuals to the important data shown

```{r}
#This visual shows the relationship between Calories burned and Total Steps tracked 

ggplot(data = daily_activity)+
  geom_point(mapping = aes(x = TotalSteps, y = Calories, color = Calories)) +
  geom_smooth(mapping = aes(x = TotalSteps, y = Calories))+ 
  ggtitle("Calories vs. Total Steps")+
  labs(title = "Calories vs. Total Steps",
       x = "Total Steps", y = "Calories")

```

```{r}
# Shows the relationship between Calories burned and Total Distance of the exercise activities 

ggplot(data = daily_activity)+
  geom_point(mapping = aes(x = TotalDistance, y = Calories, color = Calories))+
  geom_smooth(mapping = aes(x = TotalDistance, y = Calories))+
  ggtitle("Calories vs. Total Distance")+
  labs(title = "Calories vs. Total Distance",
       x = "Total Distance", y = "Calories")

```


```{r}
# Shows the relationship between the users *Time Spent in Bed* vs. *Total Steps* 

ggplot(data = combined_bars) +
  geom_bar(mapping = aes(x = TotalMinutesAsleep), stat = "count", color = "steelblue", fill = "white") +
  labs(x = "Time in Bed", y = "Steps", title = "Rest vs. Steps") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray80"),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
```


```{r}
# This graph shows the **rise** of the users steps over time 

ggplot(record_activity, aes(x = ActivityDate, y = StepTotal)) +
  geom_smooth() +
  labs(x = "Days Recorded", y = "Steps Taken", title = "Days Recorded vs. Steps Taken")

```

```{r}
# This graph shows the correlation between *longer exercise* leading to *more calories burned* 

ggplot(total_intensities, aes(x = TotalMinutes, y = Calories))+
  geom_smooth()+
  labs(x = "Total Minutes", y = "Calories Burned", title = "Intensity vs Calories Burned")

```

```{r}
# This graph shows the amount of *calories burned hourly each day* between all the participants 

ggplot(data = hourly_Calories, aes(x = ActivityHour, y = Calories))+
  geom_line(size = 1.5)+
  labs(title = "Calories Burned by Day")

```

```{r}
# Shows the amount of *steps taken hourly each day* between all the participants 

ggplot(data = hourly_steps, aes(x = ActivityHour, y = StepTotal))+
  geom_line(size = 1.5)+
  labs(title = "Hourly Steps by Day")


```

```{r}
# This graph shows the *participants weight from all of the values that were TRUE*

ggplot(data = weight_info, aes(x = Id, y = WeightPounds)) +
  geom_step(size = 2, color = "steelblue") +
  labs(title = "Participant Weight") +
  theme_minimal()

```


```{r}
# Visualizes the *weight averages for each day recorded* 

ggplot(data = weight_summary, aes(x = Date, y = AverageWeight)) +
  geom_col(fill = "steelblue") +
  labs(title = "Average Weight by Day") + 
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 week")

```

# **Act**

## **Product focus: Bellabeat app**

- By analyzing the data above I came to the conclusion that as time goes on there is more *activity* and *usage*, but towards end there seems to be a downward slope of each graph, which are both shown in the **Calories vs. Total Distance** and **Calories vs. Total Steps**. This means after a month there should be more incentives and interact-able features to keep user retention on the app

- *Sleep* and *activity* seem to be more consistent in the middle levels of the chart which is shown in the **Rest vs. Steps** chart, so I think its important for users to have at least **400-500 minutes** *6-8 hours* of sleep each day. Data like this shows the opportunity to implement systems such as *sleep trackers*, *meal plans*, *fitness guides*, *interactive profiles* and many other features to keep the users returning to the app on a consistent basis 

- Also, reviewing the *weight* of each user, we can see the users have a vast difference in weight between each other as shown in the **Participant Weight** and **Average Weight by Day** graphs. This means that each user is going to have unique goals so it's important to cater to all kinds of users by considering weight and other attributes to retain usability of the smart devices and app.

## **Conclusion**

Overall, I think Bellabeat has a great opportunity to innovate and add helpful and informative features to the app to help retain user traffic on the app so they can help reach their goals, as well as learning and helping others along the way by including an interactive interface to the app. With the Fitness tracker market contentiously growing with large growth rate potential in the future as well, there are many ways *Bellabeat* can take advantage of this opportunity to innovate by being unique compared to other competitors. As *Bellabeat* focuses on innovation they can also focus heavily on marketing to reach a larger audience of consumers that highlights the importance using the *app* to help reach personal goals and to help others along the way.

